Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2023-12-14 sections/combinatorial_semantics.tex
Pushout diagrams changes and proof for complement

Naturality is implied by $+$ being a coproduct~\cite{Frobenius}.

@@

\begin{definition}[Category $\;\catname{Csp_{D}(Hyp_{\Sigma})}$]
\label{def:cspd}
Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a \textit{full} sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
\end{definition}

@@

\[
X_{ext} \xrightarrow{f_{ext};i_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};i_1} Z_{ext}
\]
where $i_1$ is a coproduct injection.
Since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
$h = a;p_1 + g|_{Y' \setminus Y_{ext}};p_2$, $k = b;p_2 + f|_{Y \setminus Y_{ext}};p_1$.

% \[
% \begin{tikzcd}
% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
% Y'          & Y' \setminus Y_{ext} \arrow[l]
% \end{tikzcd}
% \]

@@

Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC.
More precisely, below are cospans for tensor product and symmetry.
Tensor unit is given by the initial object in $\catname{EHyp_{\Sigma}}$.

@@

@@

\begin{remark}
    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y_{ext}$.
    Consider an example in Figure~\ref{fig:e-hyp_composition_example} below.
    The pushout is depicted in Figure~\ref{fig:e-hyp_composition_example_2}.
    New interfaces are shown in Figure~\ref{fig:e-hyp_composition_example_3}.
    They are computed as shown: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken.
    Similarly for the output interface.
    In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$.
    % The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
    \update{Note that pushout for such composition always exists as the interface in the middle is descrete and nodes in the images of the middle interface are at the outermost level as well.
    Thus the span satisfies our assumptions.
    }
\end{remark}

@@

    $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v) = \varnothing$.
    And such that its sources and targets are top-level nodes.
    $h$ and $k$ are morphisms that map $X_{ext}$ and $Y_{ext}$ to these sources and targets respectively. In particular, in Figure~\ref{fig:A+B}, $h = \{c_1 \mapsto c_1, \ldots, c_n \mapsto c_n\}$ and $k = \{d_1 \mapsto d_1, \ldots, d_l \mapsto d_l\}$.
    Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
\end{definition}

@@

\begin{proposition}
    $\MdaEcospans$ is an SMC.
\begin{proof}
To show the statement above we need to first show that $\MdaEcospans$ is a proper category and then show that it has a symmetry, tensor product and the tensor unit.
In particular we need to show that the cospans that correspond to particular morphisms are mda-ones according to~\ref{def:monogamy_ehyp}.
\begin{enumerate}
    \item Id cospan is an mda-cospan. Recall the id cospan in~\ref{morphism:identity}.
    Discrete e-hypergraph $X$ is obviously directed and acyclic.
    Id-morphisms are monos.
    The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$.
    Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
    \item Composition~\ref{morphism:composition_ecospans} of two mda-cospans yields an mda-cospan.
    Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
    This also does not increase their degrees to more than 1.
    $f_{ext}$ (respectively, $g_{ext}$) is mono because its image $X$ (respectively, $Z$) is unchanged.
    The morphism $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
    Also note that $p_2$ narrowed to $B \setminus Y_{ext}$ is exactly $i_2$ and similarly for $p_1$ and $i_1$.
    Pushout is then obtained by identifying the nodes in $A + B$ that share their pre-image in $Y_{ext}$.
    These nodes are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
    $g';i_2$ is equal to $g';p_2$ because the image of $g'$ does not contain any nodes with pre-image in $Y_{ext}$.
    Next, $a$ is mono and $a;p_1$ is mono, because $p_1$ only identifies nodes with pre-image in $Y_{ext}$.
    $h = a;p_1 + g';p_2$ and hence $h$ is mono.
    Likewise, $k$ is mono.
    \item Tensor product~\ref{morphism:tensoring} of two mda-cospans is an mda-cospan.
          Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
          Hence, everything is preserved component-wise by copairing and coproduct.
    \item Symmetry~\ref{morphism:symmetry} is an mda-cospan.
          $X$ and $Y$ are both discrete as well as their coproduct and $\sigma_{X,Y}$ is obviously mono.
          Furthermore, all nodes of $X + Y$ are in the image of $\sigma_{X,Y}$ and $id_{X} + id_{Y}$.
    \item The tensor unit is discrete because it is an empty e-hypergraph and hence does not break monogamous property.
\end{enumerate}

@@

Consider an mda-cospan~\ref{mda-cospan} and an edge $e$ in $E_{A}$ such that $l(e) = \bot$. Then we take all nodes that are immediate successors of $e$. We call the nodes that have a pre-image in $X$ input nodes of $e$ and those that have a pre-image in $Y$ output nodes.
\end{definition}

% \begin{definition}[Set of inner subgraphs of an edge $e \in E_\mathcal{G}$]

% Consider an edge $e \in \mathcal{G}$ such that $l(e) = \bot$. We denote with $\mathcal{H}_{e}$ a set of inner subgraphs of the edge $e$, that is $\mathcal{H}_e \triangleq \{e' \in E_\mathcal{G} \text{ and } v' \in V_{\mathcal{G}} \; | \; \exists n,m > 0 \; . \; p^n(e') = e \text{ and } p^m(v') = e\}$. \update{And such that $\mathcal{H}_{e}$ inherits all the relations from $\mathcal{H}$ for the corresponding nodes and edges. What happens to the parent of e.g. $v$ for which $p_{V}(v) = e$?}
% \end{definition}

% add sketch of a proof for saying that a complement is an equivalence, define partition
% \begin{definition}[Instance subgraph of an edge $e \in E_\mathcal{G}$]

% Consider a partition of $\mathcal{H}_e$ according to a complement of the relation $[\#_V,\#_E]$ denoted as $[\bar{\#}_{V},\bar{\#}_{E}]$.
% Note that these relations are equivalence relations as the original relation is irreflexive. We call each $i^{th}$ subset of this partition as $i^{th}$ instance subgraph of $e$, that is, every $\mathcal{H}^i_e \subseteq \mathcal{H}_{e,\bar{\#}}$ is an instance subgraph of $e$.

@@

    Consider a hierarchical edge $e$ in $E_{A}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively.
    Then consider partitions $V_{e, in}^{\bar{\#}}$ and $V_{e,out}^{\bar{\#}}$  of the nodes according to $\bar{\#}$.
    We call $e$ well-typed if for all $v \in V_{e,in}^{\bar{\#}}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\bar{\#}}$ $|v| = |t(e)|$.
    We will call an mda-cospan in $\MdaEcospans$ well-typed if all hierarchical edges in its carrier are well-typed.
\end{definition}

@@

For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\MdaCospans$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\# = \varnothing$, and $<_{c} = \varnothing$.
It is well-typed on the nose as it has no hierarchical edges.

% \update{Two cospans $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$, $p \xrightarrow{} \mathcal{G} \xleftarrow{} q$ in $\MdaCospans$ are composable when $out^{-1}_{out}(\mathcal{G})$ and $in^{-1}_{out}(\mathcal{H})$ are isomorphic.
% The result of such composition is depicted below
% }
% \[
% \adjustbox{scale=0.75,center}{
% \begin{tikzcd}
%             &                         &                                   & k \arrow[ld] \arrow[rd] \arrow[lld] \arrow[rrd] &                                  &                         &             \\
% n \arrow[r] & \mathcal{G} \arrow[rrd] & m \arrow[l]                       &                                                 & p \arrow[r]                      & \mathcal{H} \arrow[lld] & q \arrow[l] \\
%             &                         & n \cup (p \setminus k)  \arrow[r] & \mathcal{G};\mathcal{H}                         & q \cup (m \setminus k) \arrow[l] &                         &            
% \end{tikzcd}
% }
% \]

% where $k = out^{-1}_{out}(\mathcal{G})$ and $\mathcal{G;H}$ is a pushout of $\mathcal{G} \xleftarrow{} k \xrightarrow{} \mathcal{H}$. New input interface for the composition is obtained by adding the input interface of $\mathcal{H}$ to $n$ and removing $p$ (and likewise for the output interface). $p \setminus k$ can be defined as a following pushout complement (and likewise for $m \setminus k$):

% \[
% \adjustbox{scale=0.75,center}{
% \begin{tikzcd}
% k \arrow[d] & 0 \arrow[l] \arrow[d]   \\
% p           & p \setminus k \arrow[l]
% \end{tikzcd}
% }
% \]

% then $n \cup (p \setminus k)$ is just a coproduct, since $n$ and $(p \setminus k)$ are disjoint.

% We will also use the notion of a boundary complement to preserve monogomousness when rewriting.

% \begin{definition}[Boundary complement]
%     For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
% and mono $f : L \to G$, a pushout complement as depicted below

% \begin{figure}[h!]
%     \centering
%     \includegraphics[width=0.3\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
%     \label{fig:boundary_complement}
% \end{figure}

% is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to L^{\bot}$ and $d_2 : m \to L^{\bot}$ making the above triangle commute and such that
% \[
%     n + j \xrightarrow{[d_1,c_2]} L^{\bot} \xleftarrow{[d_2,c_1]} m + i
% \]
% is a monogamous cospan
% \end{definition}

\subsection{Syntactic rewriting of PROPs}

@@

\begin{definition}[\cite{Frobenius}]
A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e. pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities.
Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as in Figure~\ref{fig:prop_rewriting}, for some rule $\langle l,r \rangle \in \mathcal{R}$    
\end{definition}

    \[
    \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
    \]
    \caption{Rewriting in $\textsf{PROP}(\Sigma)$}
    \label{fig:prop_rewriting}
\end{figure}

\begin{definition}
Given a rewriting system $\mathcal{R}$ for a $\textsf{PROP}(\Sigma)$ we define 
 a rewriting system $\mathcal{R}^{+}$ in a $\textsf{PROP}^{+}(\Sigma)$  as a set of pairs $\langle l , r \rangle$ for each pair $\langle l, r \rangle$ in $\mathcal{R}$.
However, the rewriting happens modulo semilattice enriched laws rather than SMC laws.
We define what $a \Rightarrow_{\mathcal{R}^+} b$ means for some rule $\langle l , r\rangle $ based on the structure of $a$ and $b$.

If they are decomposable as in Figure~\ref{fig:prop_rewriting} then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above ($\Rightarrow_{\mathcal{R}})$)) and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as in Figure~\ref{fig:enriched_prop_rewriting}.

\begin{figure}

    \[
    \scalebox{0.6}{
    \tikzfig{combinatorial_semantics/prop_plus_rewriting}
    }
    \]
    \caption{Rewriting in $\textsf{PROP}^{+}(\Sigma)$}
    \label{fig:enriched_prop_rewriting}
\end{figure}

% \update{

% \begin{definition}
%     Functor $\llbracket \cdot \rrbracket : \text{PROP}^{+} \to \catname{MACsp_{D}(EHyp_{\Sigma})}$.

%     We will build this functor inductively. As noted above, $\MdaCospans$ is a full subcategory of $\catname{MACsp_{D}(EHyp_{\Sigma})}$ and the acting of this functor on generators can be found in~\cite{Frobenius}. We need to define what $+$ from $\text{PROP}^{+}$ in $\catname{MACsp_{D}(EHyp_{\Sigma})}$ is. The definition could be found in Figure~\ref{fig:f+g} where well-typed mda-cospans for images of arbitrary morphisms $f, g : m \to n$ are depicted on the left and the image of $f + g$ is depicted on the right. A new edge is introduced which is the immediate predecessor of arbitrary e-hypergraphs $f$ and all edges and nodes of $f$ and $g$ are in conflict. The nodes are added to the interfaces and to the new edge to preserve well-typedness.

@@

\begin{definition}[$\ulcorner \cdot \urcorner$]

@@

    $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$ is mda-cospan in $\MdaCospans$ iff $0 \xrightarrow{} \mathcal{G} \xleftarrow{} n + m$ is an mda-cospan.
\end{proposition}

In particular, the above proposition means that by fullness of $\llbracket \cdot \rrbracket$ there exists a corresponding SMC morphism for some $0 \xrightarrow{} \mathcal{G}\xleftarrow{} n + m$.
Theorem above essentially makes two routes below. In forward direction:

@@

% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
\[\begin{tikzcd}
	{l_i+l_o} && L && {x_{ext} + y_{ext}} && R && {r_i+r_o} \\
	\\
	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
	\\
	&&&& {n_{ext}+m_{ext}}
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
	\arrow[from=1-1, to=1-3]
	\arrow[from=3-1, to=3-3]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-7, to=3-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-7]
	\arrow[from=1-9, to=1-7]
	\arrow[from=3-9, to=3-7]
\end{tikzcd}\]

The original DPOI square is shown in red.
The extra bits simply account for more complicated interfaces.
For example, the left-hand side $L$ of a rewrite rule
 $\langle L, R \rangle$ which originally was some cospan
 $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
 now is a cospan
 $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
The top half of this diagram says that the outermost interfaces
 should coincide for $L$ and $R$.
The bottom half puts the same restrictions on $G$ and $H$. 
% The extra squares on both sides say that there should exist a matching of the interfaces together with a matching of e-hypergraphs. This requirement was implicit in the case of $\catname{MACsp_{D}(Hyp_{\Sigma}}$ because the whole interfaces and the outermost interfaces coincided. Because now the interfaces can change, we need to keep track of which parts of the interfaces are subjected to changes.

@@

% Intuitively, these pushout squares mean that when we delete the occurrence of $L$ from $G$ we also remove the occurrence of $L$'s interface within $G$'s interface preserving the outermost interfaces $i + j$. The highlighted arrows form the pushout square for the original boundary complement.

% We call morphisms $i + j \to C \to G$ a boundary complement if

% \begin{enumerate}
%     \item The square above commutes and $L \xrightarrow{} G \xleftarrow{} C$ is a pushout
%     \item There exists an mda cospan $g_i \setminus (l_i \setminus i) + j \xrightarrow{} C \xleftarrow{} g_o \setminus (l_o \setminus j) + i$. \question{I am unsure about categorical notation for this set difference. Maybe it can be paraphrased in the following way?}
%     \item There exists an mda cospan $c_i + j \xrightarrow{} C \xleftarrow{} c_o + i$ where $c_i$ and $c_o$ are the following pushout complements in the category of discrete hypergraphs:
% \end{enumerate}

@@

\begin{proposition}
    The boundary complement in~\ref{def:boundary_new} when exists is unique
    \begin{proof}
        \update{TODO}
    \end{proof}
\end{proposition}

>>>>

\update{Naturality and SMC laws are implied by $+$ being a coproduct[reference?]}.

@@

\begin{definition}[Category $\;\catname{Csp_{D}(Hyp_{\Sigma})}$]
\label{def:cspd}
Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a \textit{full} sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
\update{Assuming that $D$ is a functor from PROP to $\catname{Hyp_{\Sigma}}$, $\catname{Csp_{D}(Hyp_{\Sigma})}$ is also a PROP and hence an SMC (Theorem 3.6~\cite{Frobenius}).}
\end{definition}

@@

\[
X_{ext} \xrightarrow{a_{ext};i_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};i_1} Z_{ext}
\]
where $i_1$ is a coproduct injection.
Since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
\[
    h = \langle a;p_1, g|_{Y' \setminus Y_{ext}};p_2 \rangle
\]

\[
    k = \langle b;p_2, f|_{Y \setminus Y_{ext}};p_1 \rangle
\]
\update{I've put copairings above}
% \[
% \begin{tikzcd}
% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
% Y'          & Y' \setminus Y_{ext} \arrow[l]
% \end{tikzcd}
% \]

@@

\update{Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC if $D$ is a functor from PROP to $\catname{EHyp_{\Sigma}}$.}
More precisely, below are cospans for tensor product and symmetry.
Tensor unit is given by the initial object in $\catname{EHyp_{\Sigma}}$.

@@

\question{Or shall we say that symmetry is inherited from $\catname{EHyp_{\Sigma}}$ because coproduct is commutative and associative?}
\end{definition}

@@

\begin{remark}
    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y_{ext}$.
    Consider an example in Figure~\ref{fig:e-hyp_composition_example} below.
    The pushout is depicted in Figure~\ref{fig:e-hyp_composition_example_2}.
    New interfaces are shown in Figure~\ref{fig:e-hyp_composition_example_3}.
    They are computed as shown: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken.
    Similarly for the output interface.
    In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$.
    % The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
    \update{Note that pushout for such composition always exists as the interface in the middle is descrete and nodes in the images of the middle interface are at the outermost level as well.
    Furthermore, $f_{ext};f$ and $g_{ext};g$ are monos and hence the span satisfies our assumptions.
    }
\end{remark}

@@

    $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now consistent with each other and similarly for edges and nodes from $B$.
    There is also a new hierarchical edge in $E_{C}$ that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ from $A$ and $B$ such that $[e) = \varnothing$ and $[v) = \varnothing$.
    And such that its sources and targets are top-level nodes.
    $h$ and $k$ are morphisms that map $X_{ext}$ and $Y_{ext}$ to these sources and targets respectively. In particular, in Figure~\ref{fig:A+B}, $h = \{c_1 \mapsto c_1, \ldots, c_n \mapsto c_n\}$ and $k = \{d_1 \mapsto d_1, \ldots, d_l \mapsto d_l\}$.
    Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
\end{definition}

@@

\begin{proposition}
    $\MdaEcospans$ is an SMC.
\begin{proof}
To show the statement above we need to first show that $\MdaEcospans$ is a proper category and then show that it has a symmetry, tensor product and the tensor unit.
In particular we need to show that the cospans that correspond to particular morphisms are mda-ones according to~\ref{def:monogamy_ehyp}.
\begin{enumerate}
    \item Id cospan is an mda-cospan. Recall the id cospan in~\ref{morphism:identity}.
    Discrete e-hypergraph $X$ is obviously directed and acyclic.
    Id-morphisms are monos.
    The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$.
    Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
    \item Composition~\ref{morphism:composition_ecospans} of two mda-cospans yields an mda-cospan.
    Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
    This also does not increase their degrees to more than 1.
    $a_{axt};i_{1}$ is mono because $a_{ext}$ is mono and $i_{1}$ is a coproduct injection.
    $p_1$ is mono because it only identifies $v_{1}, v_{2}$ in $V_{A}$ if there exist $y_1, y_2 \in V_{Y_{ext}}$ such that $v_{1} = f_{ext};f(y_1)$, $v_{2} = f_{ext};f(y_2)$ and there exists $v_3 \in V_{B}$ such that $v_3 = g_{ext};g(y_1)$ and $v_3 = g_{ext};g(y_2)$, but $g_{ext};g$ is mono.
    Therefore, $a;p_1$ is mono.
    $g|Y'\setminus Y_{ext}$ is mono because $g$ is and hence $g|Y'\setminus Y_{ext};p_2$ is mono and eventually $h$ is mono.
    Likewise, $k$ is mono.
    \item Tensor product~\ref{morphism:tensoring} of two mda-cospans is an mda-cospan.
          Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
          Hence, everything is preserved component-wise by coproduct.
    \item Symmetry~\ref{morphism:symmetry} is an mda-cospan.
          $X$ and $Y$ are both discrete as well as their coproduct and $\sigma_{X,Y}$ is obviously mono.
          Furthermore, all nodes of $X + Y$ are in the image of $\sigma_{X,Y}$ and $id_{X} + id_{Y}$.
    \item The tensor unit is discrete because it is an empty e-hypergraph and hence does not break monogamous property.
\end{enumerate}

@@

Consider an mda-cospan
\[
    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} C \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
\]
and an edge $e$ in $E_{C}$ such that $l(e) = \bot$.
Then we take all nodes that are immediate successors of $e$.
We call the nodes that have a pre-image in $X$ input nodes of $e$ and those that have a pre-image in $Y$ output nodes.
\end{definition}

@@

    Consider a hierarchical edge $e$ in $E_{C}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively.
    Then consider partitions $V_{e, in}^{\consistency}$ and $V_{e,out}^{\consistency}$  of the nodes according to $\consistency^{\mu}$ relation.
    We call $e$ well-typed if for all $v \in V_{e,in}^{\consistency}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\consistency}$ $|v| = |t(e)|$.
    We will call an mda-cospan in $\MdaEcospans$ well-typed if all hierarchical edges in its carrier are well-typed.
\end{definition}

@@

For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\MdaCospans$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\consistency^{\mu} = \varnothing$, and $<_{c} = \varnothing$.
It is well-typed on the nose as it has no hierarchical edges.

\subsection{DPO rewriting}

@@

% \begin{definition}[\cite{Frobenius}]
% A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e. pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities.
% Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as in Figure~\ref{fig:prop_rewriting}, for some rule $\langle l,r \rangle \in \mathcal{R}$    
% \end{definition}

% \begin{figure}[h!]

%     \[
%     \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
%     \]
%     \caption{Rewriting in $\textsf{PROP}(\Sigma)$}
%     \label{fig:prop_rewriting}
% \end{figure}

% \begin{definition}
% Given a rewriting system $\mathcal{R}$ for a $\textsf{PROP}(\Sigma)$ we define 
%  a rewriting system $\mathcal{R}^{+}$ in a $\textsf{PROP}^{+}(\Sigma)$  as a set of pairs $\langle l , r \rangle$ for each pair $\langle l, r \rangle$ in $\mathcal{R}$.
% However, the rewriting happens modulo semilattice enriched laws rather than SMC laws.
% We define what $a \Rightarrow_{\mathcal{R}^+} b$ means for some rule $\langle l , r\rangle $ based on the structure of $a$ and $b$.

% If they are decomposable as in Figure~\ref{fig:prop_rewriting} then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
% Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above ($\Rightarrow_{\mathcal{R}})$)) and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as in Figure~\ref{fig:enriched_prop_rewriting}.

% \begin{figure}

%     \[
%     \scalebox{0.6}{
%     \tikzfig{combinatorial_semantics/prop_plus_rewriting}
%     }
%     \]
%     \caption{Rewriting in $\textsf{PROP}^{+}(\Sigma)$}
%     \label{fig:enriched_prop_rewriting}
% \end{figure}

@@

\begin{definition}[\;$\ulcorner \cdot \urcorner$]

@@

% \begin{proposition}
%     $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$ is mda-cospan in $\MdaCospans$ iff $0 \xrightarrow{} \mathcal{G} \xleftarrow{} n + m$ is an mda-cospan.
% \end{proposition}

% In particular, the above proposition means that by fullness of $\llbracket \cdot \rrbracket$ there exists a corresponding SMC morphism for some $0 \xrightarrow{} \mathcal{G}\xleftarrow{} n + m$.
Theorem above essentially makes two routes below. In forward direction:

@@

% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
\[
 \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square_coloured}
 }
\]

The parts of the original DPOI square is shown in blue.
The extra bits simply account for more complicated interfaces.
For example, the left-hand side $L$ of a rewrite rule
 $\langle L, R \rangle$ which originally was some cospan
 $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
 now is a cospan
 $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
The top half of this diagram says that the outermost interfaces
 should coincide for $L$ and $R$.
The bottom half puts the same restrictions on $G$ and $H$.
For the pushout to exist we impose our assumptions~\ref{pushout:assumptions} on the correspoding arrow.

@@

@@

    The boundary complement in~\ref{def:boundary_new} when exists is unique
        \update{I will try to reinvent a similar proof from~\cite{Frobenius2} as the cited proof does not make sense to me}
        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of nodes and hyperedges.
        Hence, pushout complement should yield pushout complements for the sets of nodes and hyperedges.
        In particular,
